---
title: "Introduction to gganime"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to gganime}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)

# Fall back to code-only if the render dependencies are missing.
render_ok <- requireNamespace("gganimate", quietly = TRUE) &&
  requireNamespace("gridSVG", quietly = TRUE) &&
  requireNamespace("animejs", quietly = TRUE)
```

gganime takes a ggplot2 plot written with [gganimate](https://gganimate.com)
syntax and renders it as an animated SVG. You build the plot the same way you
would for gganimate, then call `anime()` in place of `animate()`. The output is
a self-contained HTML widget: one vector drawing of the scene with a timeline
over its elements, played in the browser.

gganime reuses gganimate's build machinery unchanged and only replaces the
final rendering step, so the [gganimate documentation](https://gganimate.com)
remains the reference for the grammar itself.

The animations below are live. Drag the scrub bar, or use the play and pause
control, to step through a frame at a time.

## Your first animation

Start from a static plot.

```{r, eval = render_ok}
library(ggplot2)
library(gganime)

p <- ggplot(iris, aes(Petal.Width, Petal.Length)) +
  geom_point()
```

Add a transition to animate it. `transition_states()` splits the data by a
discrete variable and moves the points between the resulting states.

```{r, eval = render_ok}
anim <- p +
  transition_states(Species, transition_length = 2, state_length = 1)
```

To see the animation with gganimate you would print the object or call
`animate()`. With gganime you call `anime()`:

```{r, eval = render_ok}
anime(anim, nframes = 30)
```

## Transitions

gganime supports three of gganimate's transitions.

`transition_states()`, above, animates between the levels of a discrete
variable.

`transition_time()` animates along a continuous variable, holding it to the
data range so the spacing between frames matches the spacing between values.

```{r, eval = render_ok}
aq <- ggplot(airquality, aes(Day, Temp)) +
  geom_point() +
  transition_time(Month)

anime(aq, nframes = 30)
```

`transition_reveal()` keeps earlier data on screen and reveals the rest along a
dimension, which suits a line growing over time.

```{r, eval = render_ok}
ec <- ggplot(economics, aes(date, unemploy)) +
  geom_line() +
  transition_reveal(date)

anime(ec, nframes = 40)
```

## Frame labels

A transition exposes a set of per-frame variables. Insert them into the title,
subtitle, or caption with [glue](https://glue.tidyverse.org) syntax, and the
text is swapped frame by frame in step with the geometry.

```{r, eval = render_ok}
aq_labelled <- ggplot(airquality, aes(Day, Temp)) +
  geom_point() +
  transition_time(Month) +
  labs(title = "Month: {frame_time}")

anime(aq_labelled, nframes = 30)
```

Each transition provides different variables. `transition_states()` gives
`closest_state`; `transition_time()` gives `frame_time`;
`transition_reveal()` gives `frame_along`. All three also provide `frame`,
`nframes`, and `progress`.

## Easing, entering, and exiting

`ease_aes()` sets how an aesthetic moves between values across a transition. The
default is linear; naming an aesthetic gives it its own easing.

```{r, eval = render_ok}
anime(anim + ease_aes("cubic-in-out"), nframes = 30)
```

`enter_*()` and `exit_*()` control how data that appears or leaves is drawn, so
its entrance and exit can be animated.

```{r, eval = render_ok}
anim_species <- ggplot(iris, aes(Petal.Width, Petal.Length)) +
  geom_point(aes(colour = Species), size = 2) +
  transition_states(Species, transition_length = 2, state_length = 1)

anime(anim_species + enter_fade() + exit_shrink(), nframes = 30)
```

## Several layers at once

Layers are animated independently, so you can stack them the way you would in a
static plot. Each layer keeps its own elements even when two layers draw the same
kind of shape.

```{r, eval = render_ok}
aq_layers <- ggplot(
  airquality[airquality$Month %in% c(5, 6, 7), ],
  aes(Day, Temp)
) +
  geom_area(fill = "#9ecae1") +
  geom_line(linewidth = 1, colour = "#08519c") +
  geom_point(size = 2, colour = "#a50f15") +
  labs(title = "Month {closest_state}") +
  transition_states(Month, transition_length = 2, state_length = 1)

anime(aq_layers, nframes = 30)
```

A layer that does not use the transition variable is drawn once and left in
place, which is how you add a fixed reference against the animated data.

## Leaving a trail

`shadow_mark()` keeps the marks from earlier frames on screen behind the current
one, so a scatter builds up as it advances.

```{r, eval = render_ok}
aq_shadow <- ggplot(airquality, aes(Day, Temp)) +
  geom_point() +
  transition_time(Month) +
  shadow_mark(colour = "grey70") +
  labs(title = "Month: {frame_time}")

anime(aq_shadow, nframes = 30)
```

## Controlling the render

gganime fixes the frame count and playback speed when you call `anime()`, the
same way gganimate does at `animate()`:

- `nframes` sets how many frames the animation is sampled at (default `100`).
- `fps` sets the playback speed in frames per second (default `10`), or give
  `duration` in seconds to set the speed from the total length.
- `loop` repeats the animation: `TRUE`, `FALSE`, or a number of iterations.
- `controls` shows or hides the play and scrub bar.
- `width` and `height` fix the widget size in pixels; leaving them `NULL` lets
  it fill its container.
- `precision` rounds the animated coordinates, which trims the file size.

```{r, eval = render_ok}
anime(anim, nframes = 50, duration = 4, loop = FALSE)
```

## Saving and embedding

`anime()` returns an htmlwidget, so it prints in the RStudio Viewer, knits into
R Markdown and Quarto documents, and saves to a standalone HTML file with
`htmlwidgets::saveWidget()`:

```{r, eval = FALSE}
htmlwidgets::saveWidget(anime(anim), "iris.html")
```

Loading gganime also makes a bare gganimate plot print through `anime()` at the
console, so printing `anim` gives the widget without an explicit call. Set
`options(gganime.autoprint = FALSE)` to keep gganimate's own output.

In Shiny, pair `renderGganime()` in the server with `gganimeOutput()` in the UI.

## Supported features

- Transitions: `transition_states()`, `transition_time()`,
  `transition_reveal()`.
- Geoms: points, lines and paths, bars and columns, areas and ribbons.
- Several layers in one plot, including two of the same kind, and a static layer
  beside an animated one.
- Facets: `facet_wrap()` and `facet_grid()`, with fixed or free scales.
- Coords: `coord_cartesian()`, `coord_fixed()` / `coord_equal()`, and
  `coord_flip()`.
- `enter_*()` and `exit_*()`, `ease_aes()`, and per-frame labels in the title,
  subtitle, and caption.
- `shadow_mark()`.

Anything else stops with a message naming an alternative. Current limits include
non-linear coordinate systems (`coord_polar()`, `coord_radial()`,
`coord_transform()`, `coord_sf()`), `shadow_wake()` and `shadow_trail()`, and
`view_*()` other than `view_static()`.
